\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath, amssymb}
\usepackage{array}

\title{MTH 331 — Homework \#3 Draft}
\author{Joe White}
\date{\today}

\begin{document}
\maketitle

% =========================================================
\section*{Exercise 1: Slogans, logical form, and inference rules}

\paragraph{Orville Redenbacher slogan.}
Let $P$ mean “you will like it better.” Let $Q$ mean “my name is Orville Redenbacher.”  
The slogan
\[
\text{``You will like it better or my name is not Orville Redenbacher.''}
\]
has logical form
\[
P \lor \neg Q.
\]

If the slogan is true and the speaker’s name really is Orville Redenbacher, then $Q$ is true. From $Q$ and $P\lor\neg Q$ we can conclude $P$.

\medskip
\noindent\textbf{Inference rule.}
\[
\frac{P\lor\neg Q \quad\quad Q}{P}
\]

This works because $P\lor\neg Q$ is equivalent to $Q\Rightarrow P$. With $Q$ true, modus ponens lets us conclude $P$.

\medskip
\noindent\textbf{Truth table check.}  
In the rows where both premises are true, $P$ is also true.

\begin{center}
\renewcommand{\arraystretch}{1.15}
\begin{tabular}{|c|c|c|c|c|}
\hline
$P$ & $Q$ & $\neg Q$ & $P \lor \neg Q$ & Premises $\Rightarrow P$ \\
\hline
T & T & F & T & $P$ is T \\
F & T & F & F & -- \\
T & F & T & T & -- \\
F & F & T & T & -- \\
\hline
\end{tabular}
\end{center}

\paragraph{As a conditional.}
Since $\neg Q \lor P$ is equivalent to $Q \Rightarrow P$, the slogan can also be written as
\[
Q \Rightarrow P.
\]
With $Q$ true, modus ponens again gives $P$.

\paragraph{Bumper sticker.}
“If you’re not outraged, you’re not paying attention.”  
Let $P$ mean “you are paying attention.” Let $Q$ mean “you are outraged.”  
Logical form:
\[
\neg Q \Rightarrow \neg P.
\]
The contrapositive is $P \Rightarrow Q$, meaning if you are paying attention then you are outraged.

\paragraph{Why word it this way.}
The original versions sound sharper. The Orville slogan sounds confident, and the bumper sticker is provocative. The simpler conditionals are logically the same but the given wording is more attention-grabbing.

\bigskip

% =========================================================
\section*{Exercise 3: Parity statements with three integers}

Assume $m,n,r\in\mathbb{Z}$. An integer is even if it is $2k$ for some $k$, and odd if it is $2k+1$.

\subsection*{(1) If $m,n,r$ are all odd then $m+n+r$ is odd. \,\,\textit{True.}}
Write $m=2a+1$, $n=2b+1$, $r=2c+1$. Then
\[
m+n+r = (2a+1)+(2b+1)+(2c+1) = 2(a+b+c+1)+1,
\]
which is odd.

\subsection*{(2) If $m+n+r$ is odd then $m,n,r$ are all odd. \,\,\textit{False.}}
Example: $m=2$, $n=2$, $r=1$. Then $m+n+r=5$ is odd, but $m$ and $n$ are even.

\subsection*{(3) If at least one of $m,n,r$ is odd then $m+n+r$ is odd. \,\,\textit{False.}}
Example: $m=1$, $n=1$, $r=2$. At least one number is odd, but $m+n+r=4$ is even.  
Another example: $m=1$, $n=2$, $r=3$ gives $1+2+3=6$, which is even.

\subsection*{(4) If $m+n+r$ is odd then at least one of $m,n,r$ is odd. \,\,\textit{True.}}
Proof by contrapositive. Suppose none of $m,n,r$ is odd. Then all are even: $m=2a$, $n=2b$, $r=2c$. Then
\[
m+n+r = 2(a+b+c),
\]
which is even. So if the sum is odd, not all three can be even. That means at least one must be odd.

\paragraph{Why I used these methods.}
Parts (1) and (4) work cleanly with direct proofs or a contrapositive, so that was the natural choice. 
Parts (2) and (3) are false, and the fastest way to show that is with counterexamples. 
Trying to prove them directly quickly runs into contradictions with the definitions, so a specific example is enough to disprove the claim.


\end{document}
